Stochastic Convolutional Sparse Coding

### 2018-11-08 cvpr2019/CSCmodel.tex
add paper
WH edits on full text

\section{Convolutional Sparse Coding (CSC)}
Specifically, the CSC problem has the form
\begin{equation}
\begin{split}
    \minimize{\filter,\code} & \frac{1}{2}\|\signal - \sum_{k=1}^{K} \filter_k * \code_k \|_2^2 + \lambda \sum_{k=1}^{K}\| \code_k \|_1 \\
    \text{subject to} & ~ \|\filter_k\|^2_2 \leq 1 ~~ \forall k \in \{1,\dots,K\},
\end{split}
\end{equation}
where $\signal \in \mathbb{R}^D$ is a $D$-dimensional signal or a vectorized image\footnote{In this manuscript, we work on 2D images.}, $\filter_k \in \mathbb{R}^M$ is the $k$-th dictionary, $\code_k\in \mathbb{R}^D$ is the sparse code associated with that dictionary,  $\lambda>0$ is a sparsity inducing penalty parameter, $K$ is the number of dictionary filters, and $*$ is the convolution operator. The above model will be applied to all the training images.

The modern approaches exploit Parseval's theorem and introduce two slack variables to separate the non-smooth $L_1$ penalty term and the $L_2$ constraints, making it feasible to be efficiently computed in frequency domain, and to apply splitting strategy to formulate the problem into ADMM framework, where the two subproblems (updating $\code$ and updating $\filter$) are jointly solved by coordinate descent~\cite{bristow2013fast,heide2015fast,wohlberg2016efficient}. The model and existing frequency solvers suffer from several issues:

@@

  \item We observe through experiments that the vast majority of the entries of the reconstructed sparse codes do not provide useful information about the represented image. For $K=100$, 99.5\% entries are not informative. This indicates that the subproblem for updating $\code$ solves a highly sparse LASSO problem. Transforming the problem into frequency domain imposes restriction on the use of optimization strategies for this very specific problem.

  \item While prior work shows its efficiency in solving the CSC problem in the frequency domain, this is only applicable for updating $\code$, and does not hold for updating $\filter$. The dictionary filters usually have much smaller spatial support than the dimension size of the sparse codes ($M \ll D$). However, in order to tackle the problem in the frequency domain, it requires to process the $\filter$-subproblem in the same dimensions as the sparse codes, and then project the results onto its small spatial support.
\end{itemize}

>>>>

\section{Convolutional Sparse Coding (CSC)}
The dictionary learning problem for CSC problem has the form
\begin{equation}\label{eq:CSCmodel}
\begin{split}
    \mini{\filter,\code} & \frac{1}{2}\|\signal - \sum_{k=1}^{K} \filter_k * \code_k \|_2^2 + \lambda \sum_{k=1}^{K}\| \code_k \|_1 \\
    \text{subject to} & ~ \|\filter_k\|^2_2 \leq 1 ~~ \forall k \in \{1,\dots,K\},
\end{split}
\end{equation}
where $\signal \in \mathbb{R}^D$ is a $D$-dimensional signal or a
vectorized image\footnote{In this manuscript, we work on 2D images.},
$\filter_k \in \mathbb{R}^M$ is the $k$-th dictionary, $\code_k\in
\mathbb{R}^D$ is the sparse code associated with that dictionary,
$\lambda>0$ is a sparsity inducing penalty parameter, $K$ is the
number of dictionary filters, and $*$ is the convolution operator. The
above model will be applied to all the training images
$\signal\in\mathbb{X}$.

Most recent CSC algorithms exploit Parseval's theorem and introduce
two slack variables to separate the non-smooth $L_1$ penalty term and
the $L_2$ constraints, making it feasible to efficiently compute the
latter in the frequency domain. Furthermore, the whole
Problem~\eqref{eq:CSCmodel} can be split into alternating subproblems
for updating $\code$ and $\filter$, which are jointly solved by
coordinate
descent~\cite{bristow2013fast,heide2015fast,wohlberg2016efficient}. This
approach suffers from several issues:

@@

  \item We observe through experiments that the vast majority of the
    entries of the reconstructed sparse codes do not provide useful
    information about the represented image. For $K=100$, 99.5\%
    entries are not informative. This indicates that the subproblem
    for updating $\code$ solves a highly sparse LASSO
    problem. Transforming the problem into frequency domain imposes
    restrictions on exploiting this sparsity.

  \item While prior work shows its efficiency in solving the CSC
    problem in the frequency domain, this is only applicable for
    updating $\code$, and does not hold for updating $\filter$. The
    dictionary filters usually have much smaller spatial support than
    the dimension size of the sparse codes ($M \ll D$). However, in
    order to tackle the problem in the frequency domain, it is
    necessary to process the $\filter$-subproblem over the full
    support of the sparse codes, and then project the results onto the
    much smaller spatial support of the filters.
\end{itemize}

% --- DO NOT DELETE ---
% Local Variables:
% mode: latex
% mode: flyspell
% mode: TeX-PDF
% End:

### 2018-11-08 cvpr2019/Conclusion.tex
WH edits on full text

\section{Conclusions}
In this work, we present a novel stochastic subsampling strategy for solving the CSC problem in spatial domain. This method significantly improves the runtime performance over the prior frequency-domain solvers, which applies to both batch mode and online-learning mode, as well as generating effective and proven outcomes. The proposed algorithm, for the first time, demonstrates the feasibility that tackling the CSC problem in spatial domain while still holding, or even improving the runtime efficiency. Since the subproblem of updating the code is a highly sparse LASSO, other specific optimization strategies can be applied to further accelerate the computation, for instance the idea proposed in~\cite{johnson2017stingycd}, which solves the LASSO problem by coordinate descent and skips unnecessary updates using the method of safe screening~\cite{ghaoui2012Swfe}. It would be worthy to emphasize that transforming the problem into Fourier domain cannot benefit from these kinds of speedup strategies (including the proposed one). Furthermore, instead of employing uniform subsampling strategy, alternative subsampling strategies can be studied in the future for a gain of learning efficiency, for instance an importance or correlation based sampling strategy which considers the dependency relationship between the signals and the sparse codes.

We have also shown the capability of the developed online algorithm to learn representative and meaningful over-complete dictionary from arbitrary large datasets, and the availability of the dictionary is further verified by the application of image inpainting. It can be foreseen that this capability has widespread applications in the audio and image related tasks, and higher dimensional signal processing. The source code for SBCSC and SOCSC is attached in the supplement for reference.

>>>>

\section{Conclusions}
In this work, we present a novel stochastic subsampling strategy for
solving the CSC problem in the spatial domain. This method
significantly improves the runtime performance over the prior
frequency-domain solvers, which applies to both batch mode and
online-learning mode.  The proposed algorithm, for the first time,
demonstrates the feasibility that tackling the CSC problem in spatial
domain while still holding, or even improving the runtime
efficiency. Since the subproblem of updating the code is a highly
sparse LASSO, other specific optimization strategies can be applied to
further accelerate the computation, for instance the idea proposed
in~\cite{johnson2017stingycd}, which solves the LASSO problem by
coordinate descent and skips unnecessary updates using the method of
safe screening~\cite{ghaoui2012Swfe}. It is worth emphasizing that
Fourier-domain methods cannot benefit from these kinds of speedup
strategies.

In the future, alternative subsampling strategies can be studied to
improve learning efficiency. Instead of employing uniform subsampling
strategy, for instance one could employ an importance or correlation
based sampling strategy that considers the inter-dependency between the
signals and the sparse codes.

We have also shown the capability of the developed online algorithm to
learn representative and meaningful over-complete dictionary from
arbitrary large datasets, and the availability of the dictionary is
further verified by the application of image inpainting. It can be
foreseen that this capability has widespread applications in audio
and image related tasks, and higher dimensional signal processing. The
source code for SBCSC and SOCSC is attached in the supplement for
reference.
\note{Do you really mean source code (as in full matlab code), or just
  the pseudo-code?}

% --- DO NOT DELETE ---
% Local Variables:
% mode: latex
% mode: flyspell
% mode: TeX-PDF
% End:

### 2018-11-08 cvpr2019/Introduction.tex
WH edits on full text

\section{Introduction}
Convolutional Sparse Coding (CSC) has been demonstrated to be an effective tool to learn translational invariant dictionary from a number of training signals, and shows its applications in recent researches about neural and brain information processing~\cite{jas2017learning,peter2017sparse}, as well as in a variety of image processing tasks, for instance, image inpainting~\cite{heide2015fast}, super-resolution~\cite{gu2015convolutional}, high dynamic range imaging~\cite{serrano2016convolutional}, and high-dimensional signal reconstructions~\cite{choudhury2017consensus,bibi2017high}. CSC differs from conventional sparse coding by formulating the signals as a summation of a set of convolutions on dictionary filters and sparse codes, rather than linearly adding up those filters. Alternatively, sparse coding, as a patch-based approach, learns filters from partitioned local structures, and this partition manipulation discards inherent correlations between those patches, so as to learns redundant filters (the same or similar ones with translated versions). Owing to the convolutional property, the dictionary learned from CSC achieves signal global coherence, tending to be more representative. It should be emphasized that the convolutional-based deep learning, evolving from pioneering work~\cite{lecun1998gradient,kavukcuoglu2010learning,krizhevsky2012imagenet}, has shown its extraordinary success in a broad range of high-level image understanding applications. This supervised learning based approach refers to one type of discriminative models, whereas, CSC is a kind of generative model, which, in general, is less task-specific than discriminative models.

To solve the optimization problems formulated in CSC, Zeiler et al.~\cite{zeiler2010deconvolutional} iteratively solves two subproblems (updating sparse codes and updating filters) using gradient decent in the form of convolutional operations, while performing convolution in spatial domain is computationally expensive. Recent algorithms tackles the problem by exploiting Parseval's theorem to express the spatial convolution by multiplication in the frequency domain and using Alternating Direction Method of Multipliers (ADMM)~\cite{boyd2011distributed} to separate the non-smooth terms in the optimization problem, and they show tremendous improvements over prior spatial-domain solvers with respect to running time~\cite{bristow2013fast,heide2015fast,wohlberg2016efficient}. Most of the prior work learns the dictionary filters in a batch mode, which indicates that all training signals are involved in every training iteration, and this restricts it from applying to large datasets or streaming data.

In contrast to batch mode learning, online learning~\cite{shalev2012online} is a well established strategy which processes a single or a portion (mini-batch) of the whole data at each training step, and incrementally updates model variables. Herein, the required memory and computing sources are only dependent on the sample size in every observation, independent of the training data size. It alleviates the scalability issue arisen in batch approaches, and the convergence of the algorithm was firstly analyzed using stochastic approximation tools~\cite{bottou1998online}. Bottou et al.~\cite{bousquet2008tradeoffs} further showed better generalization performance of the stochastic algorithms than standard gradient descent on large scale learning systems. Later on, online learning strategies were synergetic with sparse coding, which was then scaled up for learning dictionary from millions of training samples~\cite{mairal2009online,mairal2010online}, and for large-scale matrix factorization with an additionally introduced subsampling strategy~\cite{mensch2016dictionary}. More recently, Liu et al.~\cite{liu-2018-first} and Wang et al.~\cite{wang2018scalable} separately proposed similar online learning frameworks for the CSC model, alleviating the memory issues arise in batch-based CSC model on large datasets.

{\bfseries Contributions.} We mainly make three contributions in this work. First of all, we introduce a randomization strategy for the CSC model and solve the entire problems in spatial domain. It demonstrates that the proposed stochastic spatial-domain solver, with a reasonably selected subsampling rate, outperforms the state-of-the-art frequency-domain solvers with regard to computing efficiency. We then formulate an online-learning version of the proposed algorithm, and it shows dramatic runtime improvement over precedent online CSC methods, as well as producing comparable outcomes. At last, we demonstrate its capability to learn the over-complete dictionary from thousands of images, and analyze the effectiveness of the learned over-complete dictionary, which has not been reported or analyzed by precedent CSC work.

>>>>

\section{Introduction}
Convolutional Sparse Coding (CSC) is a method for learning {\em
  generative} models in the form of translationally invariant
dictionaries for a large variety of different training signals.  These
generative models have been shown effective for solving problems in
neural and brain information
processing~\cite{jas2017learning,peter2017sparse}, as well as in a
variety of image processing tasks, for instance, image
inpainting~\cite{heide2015fast},
super-resolution~\cite{gu2015convolutional}, high dynamic range
imaging~\cite{serrano2016convolutional}, and high-dimensional signal
reconstructions~\cite{choudhury2017consensus,bibi2017high}. CSC
differs from conventional sparse coding by formulating the signals as
the sum of a set of convolutions on dictionary filters and sparse
codes instead of patch-wise linear combinations of
filters. In traditional sparse dictionary learning, the patch
structure significantly degrades the expressiveness of the
dictionaries by introducing a strong dependency on the position of a
feature, which the convolutional nature of CSC avoids.
%% Alternatively, sparse coding, as a patch-based approach,              %%
%% learns filters from partitioned local structures, and this partition  %%
%% manipulation discards inherent correlations between those patches, so %%
%% as to learns redundant filters (the same or similar ones with         %%
%% translated versions). Owing to the convolutional property, the        %%
%% dictionary learned from CSC achieves signal global coherence, tending %%
%% to be more representative.                                            %%

Of course this convolutional approach is also a the heart of many deep
learning-based methods in the form of
CNNs~\cite{lecun1998gradient,kavukcuoglu2010learning,krizhevsky2012imagenet},
which have in recent years been extraordinarily successful for a broad
range of high-level image understanding applications. However, while
CNNs generally are used in a {\em supervised} setting and produce {\em
  discriminative}, task-specific models, CSC is {\em unsupervised} and
produces {\em generative} models that easily transfer between tasks.

To solve the optimization problems inherent to CSC, Zeiler et
al.~\cite{zeiler2010deconvolutional} iteratively solve two subproblems
(updating sparse codes and updating filters) using gradient decent in
the form of convolutional operations in the spatial domain, which is
computationally expensive. Recent algorithms tackle the problem by
exploiting Parseval's theorem to express the spatial convolution by
multiplication in the frequency domain and using proximal solver such
as Alternating Direction Method of Multipliers
(ADMM)~\cite{boyd2011distributed} to separate the linear least squares
parts terms from the non-smooth terms in the optimization
problem. These approaches show tremendous improvements over prior
spatial-domain solvers with respect to running
time~\cite{bristow2013fast,heide2015fast,wohlberg2016efficient,choudhury2017consensus}. Most
of the prior work learns the dictionary filters in a batch mode, which
indicates that all training signals are involved in every training
iteration, and this restricts it from applying to large datasets or
streaming data.

In contrast to batch mode learning, online
learning~\cite{shalev2012online} is a well established strategy which
processes a single image or a small portion (mini-batch) of the whole
data at each training step, and incrementally updates model
variables. Herein, the required memory and computing sources are only
dependent on the sample size in every observation, independent of the
training data size. It alleviates the scalability issue that arises in
batch approaches, and the convergence of the algorithm was firstly
analyzed using stochastic approximation
tools~\cite{bottou1998online}. Bottou et
al.~\cite{bousquet2008tradeoffs} further showed better generalization
performance of the stochastic algorithms than standard gradient
descent on large scale learning systems. Later on, online learning
strategies were synergetic with sparse coding, which was then scaled
up for learning dictionary from millions of training
samples~\cite{mairal2009online,mairal2010online}, and for large-scale
matrix factorization with an additionally introduced subsampling
strategy~\cite{mensch2016dictionary}. More recently, Liu et
al.~\cite{liu-2018-first} and Wang et al.~\cite{wang2018scalable}
separately proposed similar online learning frameworks for the CSC
model, alleviating the memory issues arise in batch-based CSC model on
large datasets.

{\bfseries Contributions.} We mainly make three contributions in this
work. First, we introduce a randomization strategy for the CSC
model and solve the entire problems in spatial domain. We demonstrate
that the proposed stochastic spatial-domain solver, with a reasonably
selected subsampling rate, outperforms the state-of-the-art
frequency-domain solvers with regard to computing efficiency. We then
formulate an online-learning version of the proposed algorithm, and 
show dramatic runtime improvement over current online CSC methods,
while producing comparable outcomes. Finally, we demonstrate the
capability to learn the over-complete dictionary from thousands of
images, and analyze the effectiveness of the learned over-complete
dictionary for a number of reconstruction tasks.

% --- DO NOT DELETE ---
% Local Variables:
% mode: latex
% mode: flyspell
% mode: TeX-PDF
% End:
